Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2020-09-16 paper.Rnw
simplified code

# Create a feature heatmap
plot_feature_heatmap(
  bullet_explain %>%
    mutate(
      label = as.factor(label),
      feature = fct_recode(
        feature,
        "Rough Correlation" = "rough_cor",
        "Consecutively Matching Striae" = "cms",
        "Distance" = "D",
        "Matches" = "matches",
        "Mismatches" = "mismatches",
        "Non-Consecutively Matching Striae" = "non_cms",
        "Cross Correlation Function" = "ccf",
        "Sum of Peaks" = "sum_peaks",
        "Distance Standard Deviation" = "sd_D"
      )
    ),
  facet_var = bullet_test %>%
    mutate(samesource = fct_recode(
      factor(samesource),
      "Match" = "TRUE",
      "Non-Match" = "FALSE"
    )) %>%
    mutate(samesource = fct_relevel(
      samesource, 
      "Non-Match",
      "Match"
    )) %>%
    pull(samesource) %>%
    na.omit(), 
  order_method = "PCA"
) +
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) +
  scale_color_gretchenalbrecht(palette = "last_rays",
                               discrete = TRUE,
                               reverse = TRUE) +
  theme_bw(base_family = ff, base_size = f11_fs) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.y.right = element_text(angle = 0, size = f11_fs*0.7),
    strip.text.x = element_text(size = f11_fs*0.7),
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(fill = guide_legend(nrow = 3)) +
  labs(y = "Case", color = "Complex Model Feature", fill = "Complex Model Feature")

>>>>

# Create a feature heatmap
plot_feature_heatmap(
  bullet_explain %>%
    mutate(
      label = as.factor(label),
      feature = fct_recode(
        feature,
        "Rough Correlation" = "rough_cor",
        "Consecutively Matching Striae" = "cms",
        "Distance" = "D",
        "Matches" = "matches",
        "Mismatches" = "mismatches",
        "Non-Consecutively Matching Striae" = "non_cms",
        "Cross Correlation Function" = "ccf",
        "Sum of Peaks" = "sum_peaks",
        "Distance Standard Deviation" = "sd_D"
      )
    ),
  facet_var = bullet_test %>%
    mutate(samesource = fct_recode(
      factor(samesource),
      "Match" = "TRUE",
      "Non-Match" = "FALSE"
    )) %>%
    pull(samesource), 
  order_method = "PCA"
) +
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) +
  scale_color_gretchenalbrecht(palette = "last_rays",
                               discrete = TRUE,
                               reverse = TRUE) +
  theme_bw(base_family = ff, base_size = f11_fs) +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.y.right = element_text(angle = 0, size = f11_fs*0.7),
    strip.text.x = element_text(size = f11_fs*0.7),
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(fill = guide_legend(nrow = 3)) +
  labs(y = "Case", color = "Complex Model Feature", fill = "Complex Model Feature")

### 2020-09-16 paper.tex
simplified code

{\centering \includegraphics[width=2.25in]{figure-08-1}

@@

{\centering \includegraphics[width=6.5in]{figure-10-1}

@@

We apply the trained random forest model from \citet{hare:2016} to another set of bullets from the Hamby study with 432 rows of land comparisons. We first consider a global visualization of the relationship between the random forest predictions and the model features with a parallel coordinate plot of the training data (top facet) and testing data (bottom facets) predictions (\autoref{fig:figure-10}). The majority of same source observations with random forest probabilities close to 1 have a clear pattern of corresponding feature values. The known same source observations where the random forest is wrong (does not return probabilities close to 1) have feature values that reflect those of observations where same source is known to be false.

@@

{\centering \includegraphics[width=6.25in]{figure/figure-13-1}

\end{knitrout}
\caption{Plots of LIME explanations (first and third rows) and explanation scatterplots (second and fourth rows) for  cases M and NM in the bullet test data for two tuning parameter values: 3 quantile bins (first column) and 3 equal bins (second column).  The plots provide insights into the LIME explanations and allow for the assessment of the explanation quality.}
\label{fig:figure-13}
\end{figure*}

>>>>

{\centering \includegraphics[width=2.25in]{figure-static/figure-08-1}

@@

{\centering \includegraphics[width=6.5in]{figure-static/figure-10-1}

@@

We apply the trained random forest model from \citet{hare:2016} to another set of bullets from the Hamby study with 364 rows of land comparisons. We first consider a global visualization of the relationship between the random forest predictions and the model features with a parallel coordinate plot of the training data (top facet) and testing data (bottom facets) predictions (\autoref{fig:figure-10}). The majority of same source observations with random forest probabilities close to 1 have a clear pattern of corresponding feature values. The known same source observations where the random forest is wrong (does not return probabilities close to 1) have feature values that reflect those of observations where same source is known to be false.

@@

\begin{figure*}[!thp]
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}

{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in `[[<-.data.frame`(`*tmp*`, i, value = c("{}grey50"{}, "{}darkorange"{}: replacement has 2 rows, data has 1}}\end{kframe}
\end{knitrout}
\caption{Plots of LIME explanations (first and third rows) and explanation scatterplots (second and fourth rows) for  cases M and NM in the bullet test data for two tuning parameter values: 3 quantile bins (first column) and 3 equal bins (second column).  The plots provide insights into the LIME explanations and allow for the assessment of the explanation quality.}
\label{fig:figure-13}
\end{figure*}
